### 2024-06-14 manuscript/manuscript.qmd
continue introduction

---
title: "An Evaluation of Methods for Assessing Model Fit for Bayesian Diagnostic Classification Models"
blank-lines-above-title: 2
shorttitle: "Model Fit for BDCMs"
author:
  - name: W. Jake Thompson
    corresponding: true
    orcid: 0000-0001-7339-0300
    email: jakethompson@ku.edu
    affiliations:
      - name: "University of Kansas"
        department: Accessible Teaching, Learning, and Assessment Systems (ATLAS)
        address: 1122 West Campus Road
        city: Lawrence
        region: KS
        postal-code: 66045-3101
author-note:
  blank-lines-above-author-note: 1
  disclosures:
    conflict-of-interest: The author has no conflict of interest to declare.
    financial-support: This work was supported by the Institute for Education Sciences [grant number R305D210045].
abstract: "This document is a template demonstrating the apaquarto format."
keywords: [diagnostic assessment, model fit, classification]
bibliography: ["bib/bibliography.bib", "bib/packages.bib"]
format:
  apaquarto-docx:
    floatsintext: true
  apaquarto-html:
    floatsintext: true
knitr:
  opts_chunk: 
    echo: false
    fig-dpi: 320
---

```{r}
#| label: setup
#| include: false

library(conflicted)
library(tidyverse)
library(flextable)
library(ftExtra)
library(officer)
library(knitr)
library(commonmark)
library(here)
library(glue)
library(fs)
library(english)
library(ratlas)
library(ggtext)
library(wjake)

conflicts_prefer(dplyr::filter, .quiet = TRUE)
conflicts_prefer(flextable::separate_header, .quiet = TRUE)

ratlas::set_theme(plot_margin = margin(0, 0, 0, 0))
```

Diagnostic classification models (DCMs; also known as cognitive diagnostic models [CDMs]) are a class of psychometric models where the latent traits are a set of fine-grained discrete variables [@bradshaw-dcm; @delatorre-dcm; @rupp-dcm].
That is, rather than modeling a continuous scale score, DCMs estimate a respondent's proficiency or non-proficiency on a set of predefined skills.
This results a profile of skills on which each respondent is proficient.
These profiles offer stakeholders fine-grained results that are often more useful than an overall scale score.
For example, in educational assessment, DCMs provide results that are more instructionally relevant [@zhang-2023; @lim-2024].
Similarly, applying DCMs to psychological assessment allows researchers to directly model the presence or absence of a trait, rather than attempting to apply a cut point to a continuous scale [@liu-2020; @zhang-2024].

As with all psychometric models, an evaluation of model fit is critical for users to have confidence in the inferences drawn from a DCM.
Although many methods have been proposed for evaluating model fit for DCMs, most are constrained by the way DCMs are typically estimated.
In practice, DCMs most often estimated maximum likelihood estimation [e.g., @ecpe], as this is the only process historically offered by available software [@R-GDINA; @R-CDM].
The lack of accessible alternatives to maximum likelihood estimation has in turn limited the ways in which model fit can be evaluated.
However, recent software advances have made Bayesian estimation of DCMs more accessible to practitioners [e.g., @R-blatent; @thompson-joss-2023], widening the scope of possible model fit evaluations.

In this study, we examine the efficacy of model fit indices that are available when a Bayesian estimation process is implemented.
We first provide a high-level overview of DCMs and existing model fit indices.
We then conduct a simulation study to evaluate the performance of Bayesian model fit indices and discuss the implications for the practice of diagnostic modeling.

## Diagnostic Classification Models

DCMs are confirmatory latent class models, where each class represents a particular profile of skill proficiency.
Although the latent skills, known as attributes, can be polytomous, they are most often binary.
For this discussion, we limit ourselves to binary attributes modeled with dichotomous item responses.
Using binary attributes, the number of classes is 2^*A*^, where *A* is the number of attributes measured by the assessment.
For example, an assessment measuring three attributes would have 2^3^ = 8 classes: [0,0,0], [1,0,0], [0,1,0], [0,0,1], [1,1,0], [1,0,1], [0,1,1], and [1,1,1], where 1 indicates the presence of or proficiency on the given attribute, and 0 indicates absence or non-proficiency.

In addition to the class definitions, we must also provide a Q-matrix [@tatsuoka-1983] that defines which attributes are measured by each item.
The Q-matrix has one row per item, one column per attribute.
Each cell of the Q-matrix is either a 0 (the item does not measure the attribute) or 1 (the item does measure the attribute).
Given the class definitions, the probability of respondent *r* providing a correct response is defined as

$$
P(\pmb{\text{X}}_r = x_r) = \sum_{c=1}^C\nu_c\prod_{i=1}^I\pi_{ic}^{x_{ir}}(1-\pi_{ic})^{1-x_{ir}}
$$ {#eq-dcm}

where *C* is the number of classes, *I* is the number of items, and *x~ir~* is the observed item response from respondent *r* on item *i*.
For the parameters, *&pi;~ic~* is the probability of a respondent in class *c* providing a correct response to item *i* and *&nu;~c~* is a mixing parameter that defines the base rate of membership in each class.

The definition of the &pi; parameter is determined by the particular DCM subtype that is estimated.
Each type of DCM model makes certain assumptions about how attributes interact with each other on items that measure multiple attributes.
For example, the deterministic-input, noisy-and-gate (DINA) assumes that respondents should be proficient on all attributes measured by an item in order to provide a correct response [@dina; @junker2001].
In contrast, the deterministic-input, noisy-or-gate (DINO) assumes that respondents should provide a correct response if they are proficient on any of the attributes measured by the item [@dino].

In addition to models like the DINA and DINO that make strict assumptions about attribute interactions, there are general models that make fewer assumptions and subsume the more restrictive models.
One popular general DCM is the loglinear cognitive diagnostic model (LCDM), which parameterizes &pi; similar to log-linear models [@lcdm; @lcdm-handbook].
Take as an example an item measuring two attributes.
Conditional on the attribute profile for class *c*, &alpha;~c~ = [&alpha;~1~, &alpha;~2~], the LCDM would define

$$
\text{logit}\big(P(X_{ic} = 1 | \alpha_c)\big) = \lambda_{i,0} + \lambda_{i,1,(1)}\alpha_{1} + \lambda_{i,1,(2)}\alpha_2 + \lambda_{i,2,(1,2)}\alpha_1\alpha_2
$$ {#eq-lcdm}

where &lambda;~i,0~ is an intercept and represents the log-odds of providing a correct response to item *i* when neither attribute is present.
We then have two main effects, &lambda;~i,1,(1)~ and &lambda;~i,1,(2)~, which represent the increase in the log-odds when the first or second attribute is present, respectively.
Finally, an interaction term, &lambda;~i,2,(1,2)~, represents the change in the log-odds when both both attributes are present.

By constraining the &lambda; parameters in @eq-lcdm, we can achieve models statistically equivalent to more restrictive models such as the DINA and DINO [@rupp-dcm].
If both main effects are constrained to 0, we have a model equivalent to the DINA model.
That is, a respondent is proficient on either none (&lambda;~i,0~) or both (&lambda;~i,2,(1,2)~) attributes, there is no increase in probability for being proficient on a subset of the required attributes.
The DINO model can be achieved by constraining the two main effects to be equal (i.e., &lambda;~i,1~) and constraining the interaction to be -1 &times; &lambda;~i,1~.
With these constraints, the increase in log-odds will be equal to &lambda;~i,1~ regardless of whether one or both attributes is present (i.e., the presence of any attribute is sufficient to provide a correct response).

## Model Fit for DCMs

After choosing and estimating a DCM, we must evaluate the model's performance.

### Absolute Fit

### Relative Fit

## The Current Study

# Method

To evaluate the performance of absolute and relative fit indices for DCMs, we conducted a simulation study.
In this study, we manipulated the number of assessed attributes (2 or 3), the minimum number of items measuring each attribute (5 or 7), the sample size (500 or 1,000).
These factors were chosen to represent test designs that are commonly seen in applied research [e.g., @ecpe; @dtmr].
Using a full factorial design, these factors resulted in a total of 8 test design conditions.
Within each test design condition, we also manipulated the data generating model (LCDM or DINA) and the estimated model (LCDM or DINA) in order to evaluate the performance of model fit metrics when the estimated model should and should not fit the data.
With fully crossed data generating and estimating models, there are 4 modeling conditions within each condition, resulting in a total of 32 conditions across all test designs.
We conducted 50 replications per condition.

The simulation and subsequent analyses were conducted in R version `r getRversion()` [@R-base].
All DCMs were estimated using *Stan* [version `r rstan::stan_version()`, @stan] via the measr package [@thompson-joss-2023; @R-measr], and replications were conducted on AWS EC2 instances using the portableParallelSeeds package [@R-portableParallelSeeds].
All R code for the simulation and subsequent analyses is available in a public OSF project repository.^[<https://osf.io/t5v96/>]

## Data Generation

The data generation followed the design of the data simulations used by @johnson2018 and @thompson2023 in their evaluation of reliability indices for DCMs.
In this study, the number of respondents was determined by the simulation condition.
The true attribute profile for each respondent was determined by a random draw from all possible profiles.
Additionally, each simulated assessment measured 2 or 3 attributes, which each attribute measured by at least 5 or 7 items.
The total number of items for each simulated assessment is therefore the product of the number of attributes and the minimum number of items for each attribute.
In the simulation, the Q-matrix for each simulated assessment was specified such that the first 3 items measuring each attribute are single-attribute items.
The remaining 2 or 4 items for each attribute (for the 5 and 7 item conditions, respectively) had a 50% chance of also measuring a second attribute.

Item parameter generation was dependent on the data generating model.
In conditions where data was generated from the LCDM, item parameters included item intercepts, main effects, and interactions, all of which are on the log-odds scale.
Item intercepts were drawn from a uniform distribution ranging from -3.0 to 0.6, main effects were drawn from a uniform distribution ranging from 1.0 to 5.0.
In the LCDM interaction terms are constrained to be greater than -1 times the smallest main effect.
Thus, the interaction parameters were drawn from a uniform distribution ranging from the calculated lower bound to 2.0.
In conditions where data was generated from the DINA model, item parameters include the slipping and guessing parameters, which are both on the probability scale.
For consistency with the simulation of LCDM data, parameters were generated on the log-odds scale and converted to probability values.
Guessing parameters were drawn from a uniform distribution ranging from -3.0 to 0.6, consistent with the LCDM intercepts.
The final guessing parameters were then calculated as the inverse logit of the generated parameter.
Slipping parameters were generated from a uniform distribution ranging from 1.0 to 5.0, consistent with the main effects in the LCDM.
Because the slipping parameter represents the probability of *not* providing a correct response when a respondent is proficient on the measured attributes, the final slipping parameter was calculated as 1 minus the inverse logit of the generated parameter value.

The generated attribute profiles, Q-matrix, and item parameters were then used to simulate an data set for each replication.

## Simulation Process and Analysis

After generating the data, both an LCDM and DINA model were estimated on the simulated data set.
We then calculated both absolute fit (i.e., M~2~ and raw score PPMC &chi;^2^) and relative fit (i.e., WAIC, LOO) indices for each model.
All of these indices were calculated for both estimated models in order to evaluate their performance under different conditions of known model specifications.
When data was generated from the LCDM, we expect the LCDM to show adequate model fit and be the preferred model, as the DINA model is under-specified.
On the other hand, when data was generated from the DINA model, we expect both models to show adequate absolute fit, as the LCDM subsumes the DINA model.
However, because the LCDM is over-specified in this condition, we expect the DINA model to be preferred by the relative fit indices.

For each absolute fit index, an estimated model was flagged for misfit if the *p*-value or *ppp* was less than .05 for the M~2~ and PPMC &chi;^2^, respectively.
We then used the flags to calculate the positive and negative predictive values [@altman1994; @smith2012].
The positive predictive value (PPV) is the proportion of positive results that are true positives. That is, the proportion of models where the fit index indicated poor model fit, where we expect poor model fit.
Similarly the negative predictive value (NPV) is the proportion of models where the where fit index indicated adequate model fit, and we expect adequate fit.

Finally, for relative fit, we determined the preferred model by calculating the difference between the LOO for each of the estimated models, as well as the standard error of the difference.
Using the criteria suggested by @bengio2004, if the difference between the criterion for the LCDM and DINA was greater than 2.5 times the standard error of the difference, we determined the preferred model to be the model with the lowest index value.
If the difference was less than 2.5 times the standard error, we determined the models to be equally fitting and therefore selected the more parsimonious model (i.e., the DINA model) as the preferred model.
We then calculated the proportion of replications within each condition where the LOO selected the correct model (i.e., the model that was used to generate the data).

The expected results for each combination of generating and estimating model are shown in @tbl-exp-results summarizes the expected model fit results for each of the modeling conditions.

```{r}
#| label: tbl-exp-results
#| tbl-cap: Expected model fit results
#| apa-note: |
#|   LCDM = loglinear cognitive diagnostic model;
#|   DINA = diagnostic input, noisy "and" gate.
#| ft.align: left

crossing(generate = c("LCDM", "DINA"), estimate = c("LCDM", "DINA")) |> 
  mutate(abs = case_when(generate == estimate | estimate == "LCDM" ~ "No",
                         .default = "Yes"),
         rel = generate) |> 
  rename(`Generating model` = generate,
         `Estimated model` = estimate,
         `Absolute fit flag` = abs,
         `Relative fit preference` = rel) |> 
  flextable() |>
  merge_v(j = c("Generating model", "Relative fit preference")) |> 
  hline(i = 2) |> 
  width(width = 1.2) |> 
  theme_apa() |>
  line_spacing(part = "all") |>
  padding(padding.top = 5, padding.bottom = 5)
```

# Results

```{r}
#| label: read-results
#| include: false

results <- read_rds(here("output", "sim_results.rds")) |> 
  pluck("all_reps")
```

## Absolute Model Fit

```{r}
#| label: calc-abs-fit
#| include: false

ppv <- function(x, truth, estimate) {
  x |> 
    select({{ truth }}, {{ estimate }}) |> 
    summarize(ppv = sum({{ truth }} & {{ estimate }}) / 
                sum({{ estimate }})) |> 
    pull(ppv)
}
npv <- function(x, truth, estimate) {
  x |> 
    select({{ truth }}, {{ estimate }}) |> 
    summarize(npv = sum(!{{ truth }} & !{{ estimate }}) /
                sum(!{{ estimate }})) |> 
    pull(npv)
}

abs_fit <- results |> 
  select(cond:rep, starts_with("lcdm_"), starts_with("dina_")) |> 
  pivot_longer(cols = -c(cond:rep),
               names_to = c("estimate", "metric"),
               names_pattern = c("([a-z]{4})_(.*)"),
               values_to = "value") |> 
  pivot_wider(names_from = metric, values_from = value) |> 
  relocate(estimate, .after = generate) |> 
  select(cond:rep, m2, m2_pval, ppmc_ppp) |> 
  mutate(true_flag = generate == "lcdm" & estimate == "dina",
         m2_flag = m2_pval < .05,
         ppmc_flag = ppmc_ppp < .05)

cond_abs_fit <- abs_fit |> 
  nest(flags = -c(attributes, items, sample_size)) |> 
  mutate(m2_ppv = map_dbl(flags,
                          \(x) ppv(x, truth = true_flag, estimate = m2_flag)),
         m2_npv = map_dbl(flags,
                          \(x) npv(x, truth = true_flag, estimate = m2_flag)),
         ppmc_ppv = map_dbl(flags,
                          \(x) ppv(x, truth = true_flag, estimate = ppmc_flag)),
         ppmc_npv = map_dbl(flags,
                          \(x) npv(x, truth = true_flag, estimate = ppmc_flag)))

overall_abs_fit <- abs_fit |> 
  select(cond:rep, m2, m2_pval, ppmc_ppp) |> 
  mutate(true_flag = generate == "lcdm" & estimate == "dina",
         m2_flag = m2_pval < .05,
         ppmc_flag = ppmc_ppp < .05) |> 
  nest(flags = everything()) |> 
  mutate(m2_ppv = map_dbl(flags,
                          \(x) ppv(x, truth = true_flag, estimate = m2_flag)),
         m2_npv = map_dbl(flags,
                          \(x) npv(x, truth = true_flag, estimate = m2_flag)),
         ppmc_ppv = map_dbl(flags,
                          \(x) ppv(x, truth = true_flag, estimate = ppmc_flag)),
         ppmc_npv = map_dbl(flags,
                          \(x) npv(x, truth = true_flag, estimate = ppmc_flag))) |> 
  select(-flags) |> 
  pivot_longer(cols = everything()) |> 
  deframe()
```

Across all conditions, the M~2~ statistic had a PPV of `r fmt_leading_zero(fmt_digits(overall_abs_fit["m2_ppv"], digits = 3))` and an NPV of `r fmt_leading_zero(fmt_digits(overall_abs_fit["m2_npv"], digits = 3))`.
In contrast the PPMC &chi;^2^ had a PPV of `r fmt_leading_zero(fmt_digits(overall_abs_fit["ppmc_ppv"], digits = 3))` and an NPV of `r fmt_leading_zero(fmt_digits(overall_abs_fit["ppmc_npv"], digits = 3))`.
This indicates that a negative test value for both metrics (i.e., a non-significant result) are usually true negatives.
On the other hand, a positive test result (i.e., a significant result that indicates model misfit) was a true positive only `r fmt_prop_pct(overall_abs_fit["m2_ppv"])`% of the time for the M~2~ statistic, compared to `r fmt_prop_pct(overall_abs_fit["ppmc_ppv"])`% of the time for the PPMC &chi;^2^ statistic.

When looking at the PPV and NPV values by test design condition, as show in @fig-abs-cond, we see that the NPV values for each metric are similar for all test designs.
This is consistent with the overall results, which showed similar NPVs for the M~2~ and PPMC &chi;^2^.
However, the PPV for the M~2~ is consistently lower than the PPV for the PPMC &chi;^2^.
This difference becomes more pronounced as the data set becomes larger (i.e., larger samples, more attributes).
Thus, as the sample gets larger and the test design gets more complex, the M~2~ becomes more likely to result in a false positive, indicating model misfit when there is none.
On the other hand, the PPMC &chi;^2^ demonstrated consistently high PPV values across all simulation conditions.

```{r}
#| label: fig-abs-cond
#| fig-cap: Positive and negative predictive values, by test design condition
#| out-width: "100%"

cond_abs_fit |> 
  select(-flags) |> 
  pivot_longer(cols = matches("_[pn]pv$")) |> 
  separate_wider_delim(name, delim = "_", names = c("fit_meas", "stat")) |> 
  mutate(attributes = factor(attributes, levels = c(2, 3),
                             labels = c("Two attributes", "Three attributes")),
         items = factor(items, levels = c(5, 7)),
         sample_size = factor(sample_size, levels = c(500, 1000),
                              labels = c("500", "1,000")),
         fit_meas = factor(fit_meas, levels = c("m2", "ppmc"),
                           labels = c("M<sub>2</sub>", "PPMC &chi;^2^")),
         stat = factor(stat, levels = c("ppv", "npv"),
                       labels = c("Positive predictive value",
                                  "Negative predictive value")),
         grp = paste0(attributes, items, fit_meas)) |> 
  ggplot() +
  facet_grid(cols = vars(attributes), rows = vars(stat)) +
  geom_point(aes(x = sample_size, y = value, shape = items, color = fit_meas)) +
  geom_line(aes(x = sample_size, y = value, linetype = items, color = fit_meas,
                group = grp)) +
  scale_color_okabeito() +
  expand_limits(y = c(0.6, 1)) +
  labs(x = "Sample size", y = "Predictive value",
       shape = "Items per attribute", linetype = "Items per attribute",
       color = "Absolute fit metric") +
  theme(strip.text.y = element_markdown(),
        legend.text = element_markdown())
```

## Relative Model Fit

```{r}
#| label: calc-rel-fit
#| include: false

both_fit <- abs_fit |> 
  select(cond:rep, ppmc_ppp) |> 
  pivot_wider(names_from = estimate, values_from = ppmc_ppp) |> 
  filter(lcdm > .05, dina > .05)

rel_fit <- results |> 
  semi_join(both_fit,
            join_by(cond, attributes, items, sample_size, generate, rep)) |> 
  select(cond:rep, starts_with("loo_"), starts_with("waic_")) |> 
  mutate(loo_decision = case_when(!loo_sig & generate == "dina" ~ "correct",
                                  !loo_sig & generate != "dina" ~ "wrong",
                                  loo_prefer == generate ~ "correct",
                                  loo_prefer != generate ~ "wrong"))

cond_rel_fit <- rel_fit |> 
  summarize(reps = n(),
            loo_correct = sum(loo_decision == "correct"),
            loo_wrong = sum(loo_decision == "wrong"),
            .by = c(attributes, items, sample_size, generate)) |> 
    mutate(pct_correct = loo_correct / reps)

overall_rel_fit <- rel_fit |> 
  summarize(loo_correct = sum(loo_decision == "correct"),
            loo_wrong = sum(loo_decision == "wrong")) |> 
  pivot_longer(cols = everything()) |> 
  deframe()
```

As previously discussed, evaluations of relative fit are only meaningful when the competing models have been found to have adequate absolute model fit.
Accordingly, for the relative fit results, we filtered the simulation output to only include replications where both the LCDM and DINA model showed adequate absolute model fit.
Based on the absolute fit findings in the previous section, we used the PPMC &chi;^2^ statistic to determine absolute fit.
@tbl-both-abs-fit shows the number of replications by test design condition and data generating model where both estimated model demonstrated adequate absolute fit.
As expected, both the estimated DINA and LCDM models often had adequate absolute fit when data was generated from the DINA model, as the LCDM subsumes the DINA model.
In contrast, it was much less likely that both models would show adequate absolute fit when data was generated from the LCDM model.

```{r}
#| label: tbl-both-abs-fit
#| tbl-cap: Number of replications where both models demonstrated absolute fit
#| apa-note: |
#|   LCDM = loglinear cognitive diagnostic model;
#|   DINA = diagnostic input, noisy "and" gate.
#| ft.align: left

both_fit |> 
  count(attributes, items, sample_size, generate) |> 
  pivot_wider(names_from = "generate", values_from = "n") |> 
  mutate(across(c("dina", "lcdm"), \(x) replace_na(x, 0)),
         across(everything(), as.integer)) |> 
  rename(Attributes = attributes, Items = items, `Sample size` = sample_size,
         DINA = dina, LCDM = lcdm) |> 
  flextable() |>
  add_header_row(values = c("Test designs", "Data generating model"),
                 colwidths = c(3, 2)) |> 
  width(width = 1.2) |> 
  theme_apa() |>
  line_spacing(part = "all") |>
  padding(padding.top = 5, padding.bottom = 5)
```

Overall, across all conditions, the LOO determined the correct model in `r fmt_prop_pct(overall_rel_fit["loo_correct"] / (nrow(both_fit)))`% of replications.
@fig-rel-cond shows the percentage of replications where the LOO selected the correct model by test design condition and data generating model.
Across all test design conditions, when the LCDM was used to generate the data and both the LCDM and DINA model showed adequate absolute fit, the LOO always selected correct model (i.e., the LCDM).
On the other hand, when the DINA model was used to generate the data, the LOO selected the LCDM as the preferred model in up to `r fmt_prop_pct(1 - min(cond_rel_fit$pct_correct))`% of replications (the three attribute, 7 item, and 1,000 sample size condition).
Thus, even when the DINA model was used to generate the data and the estimated DINA model showed adequate model fit, the LOO still preferred the more complex model in some situations.
However, even with a slight preference for the more complex model, the LOO still identified the correct model in greater than `r fmt_prop_pct(round_to(min(cond_rel_fit$pct_correct), .05, direction = "down"))`% of replications in all conditions.

```{r}
#| label: fig-rel-cond
#| fig-cap: Correct model selections, by test design condition
#| out-width: "100%"
#| fig-asp: 0.618

cond_rel_fit |> 
  select(attributes:reps, pct_correct) |> 
  mutate(attributes = factor(attributes, levels = c(2, 3),
                             labels = c("Two attributes", "Three attributes")),
         items = factor(items, levels = c(5, 7)),
         sample_size = factor(sample_size, levels = c(500, 1000),
                              labels = c("500", "1,000")),
         generate = factor(generate, levels = c("dina", "lcdm"),
                           labels = c("DINA", "LCDM")),
         grp = paste0(attributes, items, generate)) |> 
  ggplot() +
  facet_grid(cols = vars(attributes)) +
  geom_point(aes(x = sample_size, y = pct_correct, shape = items, color = generate)) +
  geom_line(aes(x = sample_size, y = pct_correct, linetype = items, color = generate, group = grp)) +
  scale_y_percent() +
  scale_color_okabeito() +
  expand_limits(y = c(0.6, 1)) +
  labs(x = "Sample size", y = "Repetitions With Correct Selection",
       color = "Data generating model",
       shape = "Items per attribute", linetype = "Items per attribute") +
  theme(strip.text.y = element_markdown(),
        legend.text = element_markdown())
```

# Discussion

In this study, we examined the performance absolute and relative model fit indices for Bayesian DCMs.
Overall, the findings indicate support for the use of Bayesian estimation for DCMs in order to facilitate the use of Bayesian methods for model evaluation.

For evaluating absolute model fit, we examined the M~2~ and PPMC &chi;^2^ statistics.
Across all conditions, the M~2~ statistic performed well, with results consistent with previous research evaluating the efficacy of the method [e.g., @liu2016].
However, the PPMC &chi;^2^ statistic showed comparable or improved performance in all conditions.
This was particularly true when examining the positive predictive value for each statistic (i.e., the probability that a flag is actually indicative of model misfit).
Although both the M~2~ and PPMC &chi;^2^ had similar negative predictive values, the PPMC &chi;^2^ had consistently higher positive predictive values.
Thus, when using the PPMC &chi;^2^, practitioners can have more confidence that a positive test result truly indicates model misfit.

Bayesian methods offer different methods for evaluating relative model fit than are appropriate when using a maximum likelihood estimation.
Whereas indices such as the AIC and BIC can be used when models are estimated using maximum likelihood estimation, the LOO is used for Bayesian models.
In this study, the LOO showed good performance, selecting the correct model in `r fmt_prop_pct(overall_rel_fit["loo_correct"] / (nrow(both_fit)))`% of replications.
In contrast, the AIC and BIC have been found to identify the correct model in as few as 30% of replications [@sen2017].
This means that using a Bayesian estimation process in order to access the LOO for model comparison offers a marked improvement over methods that are used with a maximum likelihood estimation.

## Limitations and Future Directions

There are some limitations to the present study.
For example, we investigated a relatively limited set of test designs.
Future research should confirm the efficacy of the PPMC &chi;^2^ and LOO under more complex designs (e.g., more attributes, more complex item structures).
Additionally, the present study assumed that a DCM was always desired, and therefore model comparisons focused on choosing between different DCM subtypes.
Future research could also consider how the LOO behaves when choosing between different psychometric models.
For example, we may compare a Bayesian DCM to an item response theory model also estimated with a Bayesian procedure.

## Conclusion

Model fit is a crucial evaluation of model performance for any psychometric model, including DCMs.
It is important not just to evaluate model fit, but also to ensure that we are using the best methods possible for the evaluation.
The present study offers evidence that the PPMC &chi;^2^ and LOO, which can only be utilized with a Bayesian estimation is used, offer improvements over existing maximum likelihood measures of model fit.
By using improved methods for model evaluation, we can greater confidence that the inferences of respondent proficiency we draw from DCMs are valid indicates of the respondents' knowledge and skills.

# References

```{r}
#| label: write-citations
#| include: false

write_package_bib <- function(x, file) {
  if (!file_exists(file)) file_create(file)

  purrr::map(
    x,
    \(p) {
      cite <- if (p == "base") {
        year <- version$version.string |> 
          str_replace_all(".*\\((.*)\\)", "\\1") |> 
          ymd() |> 
          year()

        glue::glue(
          "@manual{{R-{p},",
          "  author = {{{{R Core Team}}}},",
          "  year = {{{year}}},",
          "  title = {{{{R}}: {{A}} language and environment for statistical computing}},",
          "  version = {{Version {getRversion()}}},",
          "  type = {{Computer Software}},",
          "  publisher = {{R Foundation for Statistical Computing}},",
          "  url = {{https://www.R-project.org/}}",
          "}}",
          .sep = "\n"
        )
      } else {
        meta <- packageDescription(p)
        authors <- meta$Author |> 
          str_split_1(",\\n  ") |> 
          str_subset("\\[aut") |> 
          str_replace_all(" \\[.*$", "") |> 
          combine_words(sep = " and ", oxford_comma = FALSE)

        if (is_empty(authors)) {
          authors <- meta$Author |> 
            str_replace_all(",", " and")
        }

        year <- meta$`Date/Publication` |> 
          ymd_hms() |> 
          year()

        if (is_empty(year)) {
          year <- meta$Built |> 
            str_split_1("; ") |> 
            str_subset("UTC") |> 
            ymd_hms() |> 
            year()
        }

        glue::glue(
          "@manual{{R-{p},",
          "  author = {{{authors}}},",
          "  year = {{{year}}},",
          "  title = {{{{{p}}}: {meta$Title}}},",
          "  version = {{R package version {meta$Version}}},",
          "  type = {{Computer Software}},",
          "  publisher = {{The Comprehensive R Archive Network}},",
          "  doi = {{10.32614/CRAN.package.{p}}}",
          "}}",
          .sep = "\n"
        )
      }

    }
  ) |> 
    unlist() |> 
    xfun::write_utf8(con = file)
}

unique(c(renv::dependencies(quiet = TRUE)$Package,
         tidyverse_packages(),
         "base", "measr", "portableParallelSeeds",
         "rstan")) |> 
  write_package_bib(file = here("manuscript", "bib", "packages.bib"))

# Correct capitalization in packages
read_lines(here("manuscript", "bib", "packages.bib")) |>
  str_replace_all(" Stan", " {Stan}") |>
  str_replace_all("Bayesian", "{Bayesian}") |>
  str_replace_all("WAIC", "{WAIC}") |>
  write_lines(here("manuscript", "bib", "packages.bib"))
```

::: {#refs}
:::